---
title: "HW on Credit Card Fraud"
author: "yuexi li"
output:
  pdf_document: default
  html_document: default
---

This data set is from Kaggle. Import the data set using the **fread** function in R, and then answer the questions. 

```{r}
library(data.table) 
df_Credit<- fread('creditcard.csv')
```

Q1: Show the data elements in the file. 

```{r}
colnames(df_Credit)
dim(df_Credit)
```

Q2: Using the **lm** function, fit a regression line to show the dependence of "Class" (fraudulent=1, non-fraudulent=0) on the other variables except for "Time". (Make your dependent variable numeric if needed.) How many coefficients in your model are significant? Why? 

```{r}
library(magrittr)
x<- as.matrix(df_Credit[,c(-1,-31)])
y<- as.numeric(df_Credit$Class)
lm(y~x) %>% summary
```
Answer:
There are 29 variables have significant coefficient. 

Q3: Make sure your independent variables are in a matrix called $X$, and also assume the same is true for the dependent variable in a vector called $Y$. Enhance matrix $X$ by adding a first column of 1s. Then compute the following expression: 

$$
b = (X^\top X)^{-1} (X^\top Y)
$$

```{r echo=TRUE}
X<- cbind(1,as.matrix(df_Credit[,c(-1,-31)]))
dim(X)
b<- solve(crossprod(X,X)) %*% crossprod(X,y) %>% print
```


Q4: Compare this to the previous answer. 
The answers are the same.

Q5: What is the predicted value if the $X$ values for one credit card charge was the mean of all the $X$ variables? 

```{r}
X<- cbind(1,as.matrix(df_Credit[,c(-1,-31)]))
X_predict<- c(1, rep(mean(X),29)) 
y_predict<- t(X_predict)%*%b %>% print
```



Q6: What is the predicted value if the $X$ values for one charge is the mean of all $X$ values when Class=1? Comment.

```{r}
sub_Credit<- df_Credit[df_Credit$Class==1] 
X_class1<- c(1,rep(mean(cbind(1,as.matrix(sub_Credit[,c(-1,-31)]))),29))
y_class1<- t(X_class1)%*%b %>% print
```


The next few questions pertain to the confusion matrix shown below. 

Q7: You are given the following confusion matrix from a test sample of credit cards. The predicted values are on the rows and the true values are on the columns. Non-fraudulent charges are 0 and fraudulent charges are 1. A true positive is a fraud charge that is predicted correctly. 
              True values
            |   | 0  | 1  |        |   | 0  | 1  |
Predicted   |---|----|----|        |---|----|----|
            | 0 | 87 | 3  |        | 0 | tn | fn |
            | 1 | 4  | 15 |        | 1 | fp | tp |

Compute the statistical significance of the classifier. 

```{r}
library(magrittr)
cft_actual <- matrix(c(87,4,3,15),2) %>% print
tn <- cft_actual[1, 1] 
fn <- cft_actual[1, 2]
fp <- cft_actual[2, 1]
tp <- cft_actual[2, 2]

#Reference Dumb matrix
tn_dumb <- ((fn+tn)*(tn+fp)/(fn+tn+fp+tp)) %>% round 
fp_dumb <- ((tp+fp)*(fp+tn)/(fn+tn+fp+tp)) %>% round 
fn_dumb <- ((fn+tn)*(fn+tp)/(fn+tn+fp+tp)) %>% round 
tp_dumb <- ((tp+fn)*(fp+tp)/(fn+tn+fp+tp)) %>% round 

cft_dumb<-matrix(c(tn_dumb,fp_dumb,fn_dumb,tp_dumb),2) %>% print

cft_chisq<- (fn-fn_dumb)^2/fn_dumb + (fp-fp_dumb)^2/fp_dumb +(tp-tp_dumb)^2/tp_dumb+(tn-tn_dumb)^2/tn_dumb 
#dof=1
pchisq(cft_chisq, df=1, ncp = 0, lower.tail = FALSE) %>% print
```

Q8: What is the accuracy?

```{r}
accuracy <- ((tp + tn)/(tp + tn + fp + fn) ) %>% print
```

Q9: Compute precision, recall, and F1. 


```{r}
sprintf("percision  %f", percision <- (tp/(tp+fp)))
sprintf("recall     %f", recall <- (tp/(tp+fn)))
sprintf("F1         %f", F1<- 2/((1/recall)+(1/percision)))
```

Q10: Compute sensitivity and specificity.


```{r}
sensitivity<- recall %>% sprintf("sensitivity  %f", .) %>% print
specificity<- (tn/(tn+fp))  %>% sprintf("specificity  %f", .) %>% print
```




