---
title: "HW on Credit Card Fraud"
author: "Type your names here"
output: html_document
---

This data set is from Kaggle. Import the data set using the **fread** function in R, and then answer the questions. 

```{r}
df_Credit<- fread('creditcard.csv')
```

Q1: Show the data elements in the file. 

```{r}
colnames(df_Credit)
dim(df_Credit)
```

Q2: Using the **lm** function, fit a regression line to show the dependence of "Class" (fraudulent=1, non-fraudulent=0) on the other variables except for "Time". (Make your dependent variable numeric if needed.) How many coefficients in your model are significant? Why? 

```{r}
x_credit<- matrix(df_Credit[,c(-1,-31)])
dim(df_Credit[,c(-1,-31)])
#y<- matrix(as.numeric(df_Credit$Class))
y<- as.numeric(df_Credit$Class)
class(y)
class(x_credit)
lm(y~x_credit)

```


Q3: Make sure your independent variables are in a matrix called $X$, and also assume the same is true for the dependent variable in a vector called $Y$. Enhance matrix $X$ by adding a first column of 1s. Then compute the following expression: 

$$
b = (X^\top X)^{-1} (X^\top Y)
$$

```{r}

```

Q4: Compare this to the previous answer. 


Q5: What is the predicted value if the $X$ values for one credit card charge was the mean of all the $X$ variables? 

```{r}

```


Q6: What is the predicted value if the $X$ values for one charge is the mean of all $X$ values when Class=1? Comment.

```{r}

```


The next few questions pertain to the confusion matrix shown below. 

Q7: You are given the following confusion matrix from a test sample of credit cards. The predicted values are on the rows and the true values are on the columns. Non-fraudulent charges are 0 and fraudulent charges are 1. A true positive is a fraud charge that is predicted correctly. 

|   | 0  | 1  |
|---|----|----|
| 0 | 87 | 3  |
| 1 | 4  | 15 |

Compute the statistical significance of the classifier. 

```{r}

```

Q8: What is the accuracy?

```{r}

```

Q9: Compute precision, recall, and F1. 


```{r}

```

Q10: Compute sensitivity and specificity.


```{r}
X<- data.frame('One'=c(1,1),'Two'=c(2,2))
X^(1:1) %>% print

```




